\textbf{EECS 340: Assignment 3}

### 2020-02-21 assignment_3/article.tex
🔨 Update Question 4(b) proof with a mathematical approach.

\subsubsection{(a)}

@@

\begin{algorithm}
\caption{Partition(A, p, r)}\label{Partition}
    \begin{algorithmic}[1]
        \Procedure{Partition}{A, p, r}
        \State $q \gets p$
        \For {$i \gets p$ \textbf{to} $r$}
            \If {$\textsc{Compare-Strings}(A[i], A[r])$}
                \State $\textsc{Swap}(A[i], A[q])$ \Comment{Exchange the two elements.}
                \State $q \gets q+1$
            \EndIf
        \EndFor
        \State $\textsc{Swap}(A[q], A[r])$
        \State \Return $q$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

It is known that \textsc{QuickSort} has an average runtime of $O(n\log n)$ due to the fact that it is considered \textit{case 2} of the master theorem: as it has $n$ nodes on each level and with a depth of $\log n$, thus $O(n\log n)$. Our \textsc{QuickMiss} algorithm has a runtime of $O(n)$ due to it only calls \textsc{Partition} on either left portion or right portion to the \ilc{pivot}, but never both. Thus, on each level it will always has less than $n$ nodes, therefore forming a \textit{case 3} of the master theorem. The master theorem suggest, in \textit{case 3} the roots dominate the leaves and therefore determine the runtime of the algorithm -- as \textsc{QuickSort} has $n$ roots at the beginning, we may conclude it has a runtime of $\Theta(n)$.

>>>>

\subsubsection{(a)}\label{sec:4.a}

@@

\begin{algorithm}
\caption{Partition(A, p, r)}\label{Partition}
    \begin{algorithmic}[1]
        \Procedure{Partition}{A, p, r}
        \State $q \gets p$
        \For {$i \gets p$ \textbf{to} $r$}
            \If {$\textsc{Compare-Strings}(A[i], A[r])$}
                \State $\textsc{Swap}(A[i], A[q])$ \Comment{Exchange two elements.}
                \State $q \gets q+1$
            \EndIf
        \EndFor
        \State $\textsc{Swap}(A[q], A[r])$
        \State \Return $q$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

It is known that \textsc{QuickSort} has an average runtime of $\Theta(n\log n)$ due to the fact that it is considered \textit{case 2} of the master theorem: as it has $n$ nodes on each level and with a depth of $\log n$, thus $\Theta(n\log n)$. Our \textsc{QuickMiss} algorithm has a runtime of $\Theta(n)$ due to it only calls \textsc{Partition} on either left portion or right portion to the \ilc{pivot}, but never both (indicated in \ilc{line 4-7} of \textsc{QuickMiss}). Thus, on each level it will always have less than $n$ nodes, therefore forming a \textit{case 3} of the master theorem. The master theorem suggests, in \textit{case 3} the roots dominate the leaves and therefore determines the runtime of the algorithm -- as \textsc{QuickSort} has $n$ roots at the beginning, we may conclude it has a runtime of $\Theta(n)$.\newline

To demonstrate it mathmatically, we have \textsc{QuickSort} being $T(n) = 2T(n/2) + \Theta(n)$ where  the leading $2T(n/2)$ indicates it will recursively handle both parts of the partition. Since \textsc{QuickMiss} only needs to handle one part, we have \textsc{QuickMiss} being $T(n) = T(n/2) + \Theta(n)$. This means we have a comparsion between $n^{\log_2 1}$ and $n$, which is equivalent to $n^0 \Rightarrow 1$ and $n$. Then we may have $f(n) = \Omega(n^{\log_2 1 + \epsilon}) = \Omega(n^{0 + \epsilon})$ for $\epsilon = 1$, and we may prove that $T(n) = \Theta(n)$ for \textsc{QuickMiss}.
